\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2024-02-07 RefereeResponse/response/RefereeResponse.tex

%\documentclass[notitlepage, 11pt]{revtex4-1}
%\linespread{1.4}

\documentclass[notitlepage, 12pt]{revtex4-1}
\linespread{1.2}

\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=0.75in,rmargin=0.75in}
\usepackage{color}
\usepackage{babel}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{esint}
\usepackage[T1]{fontenc}

\usepackage{enumitem}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage{url}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{hyperref}
 \hypersetup{colorlinks=true,citecolor=blue,linkcolor=blue}
\renewcommand{\bibfont}{\footnotesize}
\usepackage{fancyheadings}
%\pagestyle{fancy}
\renewcommand{\headrulewidth}{12pt}

\makeatother

\usepackage{babel}
\begin{document}

\title{Response to Referee: JHEP\_031P\_1223}

\date{\today}

\maketitle

We would like to begin by sincerely thanking the referee for their careful reading of the manuscript.
We believe their thoughtful comments and suggestions have produced a better paper.

The referee pointed out that the manuscript contained insufficient context for the two-loop results.
Specifically, the referee noted that a comparison with $\mathcal{N}=4$ super-Yang Mills (sYM) was absent.
We could not agree more and we apologize for the oversight.
Below is summary of the modifications to the manuscript.
\begin{itemize}
\item A small comment has been added to the introduction at the top of page 2 reminding readers that color-dual representations of sYM have been found to high loop order.
\item At the bottom of page 2, when stating that there is no local 4pt 2-loop numerator for pure YM we remind readers that such a numerator exists for sYM.
\item At the end of the introduction at the top of page 3 we clarify that the bowtie cut failure is specific to pure YM.
We go on to speculate that this is because sYM and NLSM are simpler theories, at least as far as color-kinematics is concerned.
\item Although the referee suggested adding comments to the introduction and section 4, we have included one additional modification.
Section 2.3 deals with ansatz locality so this seemed like a natural place to mention that the sYM ans\"atze incorporate locality as a key simplifying assumption.
At the bottom of page 7 and the top of page 8, we emphasize that the sYM numerators through four loops are local.
Even though the double copy has been performed for sYM at five loops, local numerators have not been found.
We comment in footnote 4 that the full set of local numerators has never been fully explored for sYM at 4pt 5-loop.
\item The first paragraph in section 4 closes by putting our results in the context of sYM, specifically by reminding readers of the state of the art in that theory.
\item Footnote 10 at the beginning of section 4.3 reminds readers again of the state of the art in sYM.
\item After stating the bowtie failure condition for pure YM in (4.9) we state that NLSM and sYM do not suffer from this failure, likely because they are simpler theories in some sense.
\end{itemize}

We would like to thank the referees again for their carefully considered feedback.
Their suggestions have improved the readability of the manuscript.
We believe all of their comments have been addressed and the the paper should be ready for publication.

\

Best regards,

Alex Edison, James Mangan, and Nic Pavao

\end{document}

### 2024-02-07 RefereeResponse/paper/2loopNLSM.tex
Referee response v1

Despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv}, color-kinematics remains a conjecture at loop
level. Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies to advance our
understanding of the kinematic algebra at the multi-loop level.

@@

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories.  However, in sharp contrast, we find that Yang-Mills does
\emph{not} permit a globally color-dual representation, even when
considering the most general polynomial ansatz of Lorentz covariant
kinematics.\footnote{Globally color-dual local numerators for two-loop
  pure YM have been studied before but only in the context of a
  limited ansatz subject to certain loop power counting constraints
  \cite{Bern:2015ooa}.}  Concretely, the failure point can be
pinpointed to a conflict between the ``bowtie'' cut and the
following Jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions}
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated since Feynman
rules produce polynomial numerators. Indeed, the most natural way to guarantee locality and avoid spurious poles is to require the off-shell numerators to be polynomial functions of
kinematics. However, insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  While these
assumptions are well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.  However, simply requiring on-shell locality may suffice.
This alternative will be discussed in \cref{nonLocalScattering}.

@@

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
one-loop numerators at any multiplicity for pions and related
theories, but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the two-loop no-go statement
is only for \emph{one particular representation of the theory}, so it does
not completely preclude the existence of a two-loop color-dual
integrand.  We thus turn to the color-dual bootstrap method
described in \cref{sec:bootstrap} to construct a color-dual
representation of two-loop four-point NLSM.  Because of the similarity
of the problem setup, we will also use the opportunity to revisit the
construction of a color-dual representation of pure YM, extending the
search space beyond what was covered in Ref.~\cite{Bern:2015ooa} to include the most general local ansatz.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the
Jacobi basis graphs and building the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown in
\cref{fig:ym-spanning}, but we choose to work within the framework of
the method of maximal cuts \cite{Bern:2007ct} in order to identify the
simplest cut that is in tension with the kinematic Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14
non-pathological cubic diagrams shown above in \cref{fig:MaxCuts}.
Doing so, we are left with 596 total parameters in the ansatz.

@@

In fact, we can make the failure extremely precise. First, start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}, which includes
the double-box, crossed-box, and penta-triangle. Then \emph{without} defining the crossed-box in terms of the other two diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop) for each of the three diagrams.  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2,761 free parameters on the double-box, 2,576 free parameters on the
crossed-box, and 5,040 free parameters on the penta-triangle.  Finally
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only receives contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied. In other words, the constraints imposed by the single Jacobi relation and the three sets of diagram symmetries conspire in a way that is inconsistent with the ``bowtie'' cut. The contrapositive to this surprising result can be summarized in the following diagrammatic statement, which prioritizes the physical unitarity cut:
\begin{eBox}
\begin{equation}\label{chefsKiss}
   \scalebox{0.9}{\NMCutC} \!\!
  \Rightarrow \!\!
  \scalebox{0.85}{\JacobiOneDoubleBox}
  +  \scalebox{0.85}{\JacobiOneCrossedBox}
  + \scalebox{0.85}{\JacobiOnePentaTriangle} \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: off-shell locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.

>>>>

Despite the tremendous success in leveraging this duality to
compute gauge and gravity observables to high orders in perturbation
theory, notably $\mathcal{N}=4$ super-Yang Mills (sYM) at five loops \cite{FiveLoopN4, GeneralizedDoubleCopyFiveLoops, Bern:2018jmv}, color-kinematics remains a conjecture at loop
level. %Referee edit
Indeed, there are many examples in the literature where
identifying color-dual representations beyond one-loop has posed a
formidable challenge
\cite{Mogull:2015adi, Johansson:2017bfl, KalinN2TwoLoop, Bern:2015ooa}. In this work, we use
the nonlinear sigma model (NLSM) and Yang-Mills, two theories proven
to permit color-dual representations at tree-level
\cite{Feng:2010my,Cachazo:2014xea}, as case studies to advance our
understanding of the kinematic algebra at the multi-loop level.

@@

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric
theories.
However, in sharp contrast to both sYM \cite{BRY} and NLSM, we find that Yang-Mills does %Referee edit.  Am I citing the right paper?
\emph{not} permit a globally color-dual two-loop representation, even when
considering the most general polynomial ansatz of Lorentz covariant
kinematics.\footnote{Globally color-dual local numerators for two-loop
  pure YM have been studied before but only in the context of a
  limited ansatz subject to certain loop power counting constraints
  \cite{Bern:2015ooa}.}
  Concretely, the failure point can be
pinpointed to a conflict between the ``bowtie'' cut and the
following Jacobi triple:
\begin{equation}
\label{eq:IntroBowtieFailure}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
This failure is specific to Yang-Mills and does not occur for its maximally supersymmetric counterpart or for NLSM, indicating that pure Yang-Mills is a more complicated theory in some sense. %Referee edit
We comment on the implications of \cref{eq:IntroBowtieFailure} in \cref{conclusions}
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

@@

The kinematic numerators must be $D$-dimensional, Lorentz invariant,
local (polynomial) functions of the graph kinematics with the correct
power counting and external states. Locality is physically motivated since Feynman
rules produce polynomial numerators.
Indeed, the most natural way to guarantee locality and avoid spurious poles is to require the off-shell numerators to be polynomial functions of kinematics.
However, insisting on a local numerator is
partly a matter of convenience as the ansatz for a rational function
would quickly grow out of control without a guiding principle for what
terms to include.  %Referee edit
Historically, locality has also proved a highly successful tool for simplifying kinematic numerators. %Alex, can you add more relevant references to the next two sentences?
For example, the four-point function for $\mathcal{N}=4$ super-Yang Mills admits local color-dual numerators through four loops \cite{FivePointN4BCJ}.\footnote{At five loops, the most general local ansatz has not been explored, but the double copy has been performed at five loops using alternative methods \cite{FivePointN4BCJ}.}
While locality is well motivated from a physical and complexity
standpoint, the literature has featured numerators that violate
locality, modify naive power counting through the inclusion of higher
spin modes, and break manifest Lorentz invariance \cite{Square,
  WeinzierlBCJLagrangian, Mogull:2015adi, FivePointN4BCJ,
  Johansson:2017bfl, Ben-Shahar:2022ixa, Cheung:2016prv,
  Cheung:2021zvb}.  These developments suggest that relaxing locality
is a viable alternative when the condition proves too stringent for an
ansatz.  However, simply requiring on-shell locality may suffice.
This alternative will be discussed in \cref{nonLocalScattering}.

@@

\section{Two-loop four-point bootstrap}
\label{2loopBoot}
As we have just seen, it is possible to coerce tree numerators into
one-loop numerators at any multiplicity for pions and related
theories, but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the two-loop no-go statement
is only for \emph{one particular representation of the theory}, so it does
not completely preclude the existence of a two-loop color-dual
integrand.  We thus turn to the color-dual bootstrap method
described in \cref{sec:bootstrap} to construct a color-dual
representation of two-loop four-point NLSM.  Because of the similarity
of the problem setup, we will also use the opportunity to revisit the
construction of a color-dual representation of pure YM, extending the
search space beyond what was covered in Ref.~\cite{Bern:2015ooa} to include the most general local ansatz.
Even though this work is concerned with non-supersymmetric theories, recall that \ansatze{} techniques have been very successful for sYM.  %Referee edit
In particular, local color-dual representations have been found through 4pt 4-loop \cite{}. %What paper should we cite?
While a local representation has not been found at five loops, the double copy has still be performed \cite{FivePointN4BCJ}.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.\footnote{As mentioned at the beginning of this section, \emph{supersymmetric} Yang-Mills is known to have a local color-dual representation at least through four loops \cite{BRY}.  A local but restricted ansatz fails at five loops, but the most general local ansatz has not been explored \cite{FivePointN4BCJ}.} %Referee edits.  What papers should we cite?
We follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the
Jacobi basis graphs and building the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot products instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form
\begin{equation}
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  N^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  Without any power counting restrictions
imposed, both the double-box and the penta-triangle have an ansatz
with 10,010 terms each.  Imposing diagram automorphisms and
maximal-cut gauge invariance on the two basis diagrams, we reduce the
number of terms to 2,235 for the double-box, and 4,133 for the
penta-triangle.  The minimal set of spanning physical cuts is shown in
\cref{fig:ym-spanning}, but we choose to work within the framework of
the method of maximal cuts \cite{Bern:2007ct} in order to identify the
simplest cut that is in tension with the kinematic Jacobi relations.
Maximal cuts and symmetries are then imposed on the 14
non-pathological cubic diagrams shown above in \cref{fig:MaxCuts}.
Doing so, we are left with 596 total parameters in the ansatz.

@@

In fact, we can make the failure extremely precise. First, start with the
diagrams shown in the Jacobi relation from \cref{eq:TwoLoopJacobiOne}, which includes
the double-box, crossed-box, and penta-triangle. Then \emph{without} defining the crossed-box in terms of the other two diagrams, write down the most generic parity-even local ansatz
involving four polarizations and six momenta (any combination of
external or loop) for each of the three diagrams.  Each ansatz will have 10,010 terms initially.  Next
impose the symmetry constraints on each of the three diagrams, leaving
2,761 free parameters on the double-box, 2,576 free parameters on the
crossed-box, and 5,040 free parameters on the penta-triangle.  Finally
impose the ``bowtie'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only receives contributions
from the double-box
\begin{equation}
   \NMCutC
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  N \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  N\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied. In other words, the constraints imposed by the single Jacobi relation and the three sets of diagram symmetries conspire in a way that is inconsistent with the ``bowtie'' cut. The contrapositive to this surprising result can be summarized in the following diagrammatic statement, which prioritizes the physical unitarity cut:
\begin{eBox}
\begin{equation}\label{chefsKiss}
   \scalebox{0.9}{\NMCutC} \!\!
  \Rightarrow \!\!
  \scalebox{0.85}{\JacobiOneDoubleBox}
  +  \scalebox{0.85}{\JacobiOneCrossedBox}
  + \scalebox{0.85}{\JacobiOnePentaTriangle} \neq 0 \,.
\end{equation}
\end{eBox}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang-Mills representation: off-shell locality,
symmetry, and the kinematic Jacobi relations for just three diagrams
are incompatible with the ``bowtie'' cut.
Note that NLSM and sYM do not suffer from this failure, suggesting that they are more privileged theories in some sense.%Referee edit
